Matching Bullets

### 2015-12-09 imaging-paper.Rnw
trying to make figures show up closer to their text

\usepackage[USenglish]{babel}
\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
\usepackage{float}

@@

@@

\begin{figure}
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='\\linewidth'>>=
library(RColorBrewer)
library(ggplot2)
library(scales)
library(dplyr)
library(x3pr)
library(x3prplus)
library(grid)
library(gridExtra)
library(zoo)
library(tidyr)
library(rpart)
library(rpart.plot)
library(xtable)

@@

\begin{figure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residuala}Residual structure of a surface crosscut at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residualb} Residual structure of a surface crosscut at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of  circular fits at two different cross-section. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate. }
\end{figure}

@@

\begin{figure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm. For a fixed crosscut ($x$ held constant at 243.75 $\mu m$) of surface measurements for land 5 of bullet 1 are plotted across different values of $y$.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
br121_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

@@

\begin{figure}
<<side-by-side, echo=FALSE, eval=FALSE>>=
paths <- file.path(datadir, dir(datadir))

@@

\includegraphics[width=\columnwidth]{images/side-by-side-manual-match.png}
\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
\end{figure}
\begin{figure}
\includegraphics[width=\columnwidth]{images/matchup-rgl.png}
\caption{\label{fig:manualmatch-rgl}\hh{Different view of the manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
\end{figure}

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}
\centering
\begin{subfigure}[t]{\linewidth}
\caption{Crosscuts at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of Loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position="none") + ylab("Residuals of Loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}
<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '\\linewidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation between the two crosscuts shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the crosscuts. Figure~\ref{fig:crosscutX} shows the lag-shifted crosscuts.}
\end{figure}

\begin{figure}
\centering
    \begin{subfigure}[t]{\linewidth}
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bullet~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Using a rolling median peaks and grooves are identified for each crosscut. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  lof <- processBullets(paths = images[c(5,7)], x = 100)
  lof <- bulletSmooth(lof)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
  ylim(c(-6,6.5)) + theme_bw() +
  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, linetype=bullet)) +
  scale_linetype_discrete("") +
  scale_colour_manual("", values=cols) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) +
  xlab("y")
@
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between crosscuts (c).}
\end{figure}

\subsection{Varying crosscuts}
The crosscut at which a comparison is done matters -- as crosscuts are further apart, more pronounced differences between the crosscuts show up. For bullet land 1-5 from barrel 1 crosscuts at intervals of 50$\mu m$ between 150$\mu m$  and 450$\mu m$  are compared to a crosscut at 100$\mu m$. The initially almost perfect match (consolidated by a CMS of fourteen) between the  crosscuts deteriorates with increasing distance to non-matches at a crosscut of $x = 300\mu m$ or higher. Only if the crosscut is within 150$\mu m$ do we get a good visual match. Given that alignment in the microscope is not perfect, we should take this into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}
<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100, 450, by = 50)
lof <- processBullets(paths[5], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

lop <- lapply(reslist, function(res) {
ggplot() +
  theme_bw() + 
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
  geom_line(aes(x = y, y = l30, linetype=bullet),   data = res$bullets, alpha=0.6) +
  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
  scale_linetype_discrete("") +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylim(c(-6,6)) +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
    ylab("")
})

grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]], 
             lop[[7]],
             ncol = 1)
@
\caption{\label{fig:crosscuts}Overview of the variations in the surface along different crosscuts. Crosscut at $x = 100$ is compared to crosscuts every 50$\mu m$. With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher. }
\end{figure}

%\hh{XXX Does this give us a way to automatically check for the bottom of the bullet? Attempt below} 
By comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section, indicative of a break-off, or in a stable section, where we can expect to find matches reasonably well. In the approach here, we keep increasing the crosscut level until we find a section with a stable pattern. This process is shown  in Figure~\ref{fig:crosscuts2} at the example of land 1-1 from barrel 3.
\hh{`Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9. }
\begin{figure}
<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100,225, by=25)
lof <- processBullets(paths[37], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

lop <- lapply(reslist, function(res) {
ggplot() +
  theme_bw() + 
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
  geom_line(aes(x = y, y = l30, linetype = bullet),   data = res$bullets) +
  scale_linetype_discrete("") +
#  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylim(c(-6,6)) + ylab("") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
})

@@

\subsection{Automated matching algorithm}
\hh{The automated matching routine is then as follows:
Identify a stable region for each bullet and pick the lowest crosscut level in this region (assumption here: lower crosscut levels show better expression of individual marks XXX need to still check)

@@

\begin{figure}
<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='\\linewidth', warning=FALSE>>=

@@

grid.arrange(list_of_matches[[1]], list_of_matches[[2]], list_of_matches[[3]], 
             list_of_matches[[4]], list_of_matches[[5]], list_of_matches[[6]],
             ncol = 1)
@
\caption{\label{fig:barrel2}All lands from the first bullet shot from barrel 2 are compared to the first land of the second bullet shot from barrel 2. The largest number of CMS is eleven, observed with land six of the first bullet. }
\end{figure}

@@

\begin{figure}
\begin{subfigure}[t]{.49\textwidth}
\caption{Maximal number of CMS between unknown Bullet F land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
<<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
load("data-35-25/unkn27.RData")
cmsdist <- sapply(reslist, function(x) x$maxCMS)

@@

Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 29 pairs of lands with at least sixteen CMS in common are matches. For smaller number of CMS, the percentage of false positives quickly increases. Taking cross-correlation into account (see Figure~\ref{fig:ccfs-match}) increases the number of matches. A cross-correlation of above 0.75 in the Hamby study is indicative of a match. There are 97 land-to-land comparisons with a cross-correlation of above 0.75.   
\begin{figure}
\begin{subfigure}[t]{\linewidth}
\caption{Distribution of maximal CMS.}
<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
bstats <- read.csv("data-35-25/bullet-stats.csv")
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms))) + theme_bw() +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{All land-to-land comparisons a maximum number of CMS of sixteen or more are matches.}
<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms), fill=match), position="fill") +
  theme(legend.position="bottom") + 
  scale_fill_brewer(palette="Paired") +
  xlab("maximum CMS")
@
\end{subfigure}
\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
\end{figure}

>>>>

\usepackage[USenglish]{babel}
\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
\usepackage{float}
\usepackage{afterpage}

@@

    \renewcommand{\topfraction}{0.99}	% max fraction of floats at top
    \renewcommand{\bottomfraction}{0.99}	% max fraction of floats at bottom

@@

\begin{figure}[hbtp]
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='\\linewidth'>>=
library(RColorBrewer)
library(ggplot2)
library(scales)
library(dplyr)
library(x3pr)
library(x3prplus)
library(grid)
library(gridExtra)
library(zoo)
library(tidyr)
library(rpart)
library(rpart.plot)
library(xtable)

@@

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residuala}Residual structure of a surface crosscut at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residualb} Residual structure of a surface crosscut at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of  circular fits at two different cross-section. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate. }
\end{figure}

@@

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm. For a fixed crosscut ($x$ held constant at 243.75 $\mu m$) of surface measurements for land 5 of bullet 1 are plotted across different values of $y$.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
br121_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

@@

\begin{figure}[hbtp]
<<side-by-side, echo=FALSE, eval=FALSE>>=
paths <- file.path(datadir, dir(datadir))

@@

\includegraphics[width=\columnwidth]{images/side-by-side-manual-match.png}
\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
\end{figure}
\begin{figure}[hbtp]
\includegraphics[width=\columnwidth]{images/matchup-rgl.png}
\caption{\label{fig:manualmatch-rgl}\hh{Different view of the manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
\end{figure}

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}[hbtp]
\centering
\begin{subfigure}[t]{\linewidth}
\caption{Crosscuts at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of Loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position="none") + ylab("Residuals of Loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}[hbtp]
<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '\\linewidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation between the two crosscuts shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the crosscuts. Figure~\ref{fig:crosscutX} shows the lag-shifted crosscuts.}
\end{figure}

\subsection{Varying crosscuts}
The crosscut at which a comparison is done matters -- as crosscuts are further apart, more pronounced differences between the crosscuts show up. For bullet land 1-5 from barrel 1 crosscuts at intervals of 50$\mu m$ between 150$\mu m$  and 450$\mu m$  are compared to a crosscut at 100$\mu m$. The initially almost perfect match (consolidated by a CMS of fourteen) between the  crosscuts deteriorates with increasing distance to non-matches at a crosscut of $x = 300\mu m$ or higher. Only if the crosscut is within 150$\mu m$ do we get a good visual match. Given that alignment in the microscope is not perfect, we should take this into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100, 450, by = 50)
lof <- processBullets(paths[5], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

lop <- lapply(reslist, function(res) {
ggplot() +
  theme_bw() + 
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
  geom_line(aes(x = y, y = l30, linetype=bullet),   data = res$bullets, alpha=0.6) +
  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
  scale_linetype_discrete("") +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylim(c(-6,6)) +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
    ylab("") + xlab("")
})

grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]], 
             lop[[7]],
             ncol = 1)
@
\caption{\label{fig:crosscuts}Overview of the variations in the surface along different crosscuts. Crosscut at $x = 100$ is compared to crosscuts every 50$\mu m$. With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher. }
\end{figure}
\afterpage{\clearpage}

%\hh{XXX Does this give us a way to automatically check for the bottom of the bullet? Attempt below} 
By comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section, indicative of a break-off, or in a stable section, where we can expect to find matches reasonably well. In the approach here, we keep increasing the crosscut level until we find a section with a stable pattern. This process is shown  in Figure~\ref{fig:crosscuts2} at the example of land 1-1 from barrel 3.
\hh{`Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9. }
\begin{figure}[hbtp]
<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100,225, by=25)
lof <- processBullets(paths[37], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

lop <- lapply(reslist, function(res) {
ggplot() +
  theme_bw() + 
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
  geom_line(aes(x = y, y = l30, linetype = bullet),   data = res$bullets) +
  scale_linetype_discrete("") +
#  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylim(c(-6,6)) + ylab("") + xlab("") + 
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
})

@@

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\linewidth}
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bullet~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Using a rolling median peaks and grooves are identified for each crosscut. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  lof <- processBullets(paths = images[c(5,7)], x = 100)
  lof <- bulletSmooth(lof)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
  ylim(c(-6,6.5)) + theme_bw() +
  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, linetype=bullet)) +
  scale_linetype_discrete("") +
  scale_colour_manual("", values=cols) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) +
  xlab("y")
@
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between crosscuts (c).}
\end{figure}
\afterpage{\clearpage}

\subsection{Automated matching algorithm}

\hh{The automated matching routine is then as follows:
Identify a stable region for each bullet and pick the lowest crosscut level in this region (assumption here: lower crosscut levels show better expression of individual marks XXX need to still check)

@@

\begin{figure}[hbtp]
<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='\\linewidth', warning=FALSE>>=

@@

grid.arrange(list_of_matches[[1]], list_of_matches[[2]], list_of_matches[[3]], 
             list_of_matches[[4]], list_of_matches[[5]], list_of_matches[[6]],
             ncol = 1)
@
\caption{\label{fig:barrel2}All lands from the first bullet shot from barrel 2 are compared to the first land of the second bullet shot from barrel 2. The largest number of CMS is eleven, observed with land six of the first bullet. }
\end{figure}
\afterpage{clearpage}

@@

\begin{figure}[hbtp]
\begin{subfigure}[t]{.49\textwidth}
\caption{Maximal number of CMS between unknown Bullet F land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
<<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
load("data-35-25/unkn27.RData")
cmsdist <- sapply(reslist, function(x) x$maxCMS)

@@

Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 29 pairs of lands with at least sixteen CMS in common are matches. For smaller number of CMS, the percentage of false positives quickly increases. Taking cross-correlation into account (see Figure~\ref{fig:ccfs-match}) increases the number of matches. A cross-correlation of above 0.75 in the Hamby study is indicative of a match. There are 97 land-to-land comparisons with a cross-correlation of above 0.75.   
%
\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{Distribution of maximal CMS.}
<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
bstats <- read.csv("data-35-25/bullet-stats.csv")
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms))) + theme_bw() +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{All land-to-land comparisons a maximum number of CMS of sixteen or more are matches.}
<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms), fill=match), position="fill") +
  theme(legend.position="bottom") + 
  scale_fill_brewer(palette="Paired") +
  xlab("maximum CMS")
@
\end{subfigure}
\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
\end{figure}
